Visualising metagenome differences

### 2023-11-23 analysis/metagenome.Rmd
another try

@@

@@

>>>>

### GTDB order-level stats
Some stats. How many MAGs on order level?

```{r}
amph_pos_tax |> filter(order != 'o__') |> count(order) |> nrow()
```
```{r}
amph_pos_tax |> filter(order != 'o__') |> count(order) |> summarise(total = sum(n))
```
MAGs per species:

```{r}
amph_pos_tax |> group_by(Sample_species) |> filter(order != 'o__') |> count(order) |> summarise(total = sum(n))

How many of these are unique to either species?
```{r}
in_amph <- amph_pos_tax |> filter(Sample_species == 'Amphibolis antarctica') |> filter(order != 'o__') |> count(order)
in_posi <- amph_pos_tax |> filter(Sample_species == 'Posidonia australis') |> filter(order != 'o__') |> count(order)
```

Unique orders in Amphibolis:
```{r}
in_amph |> nrow()
```

Unique orders in Posidonia:
```{r}
in_posi |> nrow()
```

In Amphibolis, not in Posidonia:
```{r}
setdiff(in_amph$order, in_posi$order) |> length()
```
```{r}
in_amph |> filter(order %in% setdiff(in_amph$order, in_posi$order)) |> arrange(desc(n)) |> head(5)
```

In Posidonia, not in Amphibolis:

```{r}
setdiff(in_posi$order, in_amph$order) |> length()
```
```{r}
in_posi |> filter(order %in% setdiff(in_posi$order, in_amph$order)) |> arrange(desc(n))
```

@@

### CAT order-level stats

Some more stats here too!

```{r}
amph_pos_tax_cat |> filter(!is.na(Order)) |> separate(Order, into=c('Order', 'Order_score'), sep=':') |> count(Order) |> nrow()
```
```{r}
amph_pos_tax_cat |> filter(!is.na(Order)) |> separate(Order, into=c('Order', 'Order_score'), sep=':') |> count(Order) |> summarise(total = sum(n))
```
MAGs per species:

```{r}
amph_pos_tax_cat |> group_by(Sample_species)|> filter(!is.na(Order)) |> separate(Order, into=c('Order', 'Order_score'), sep=':') |> count(Order) |> summarise(total = sum(n))
```

How many of these are unique to either species?
```{r}
in_amph <- amph_pos_tax_cat |> filter(Sample_species == 'Amphibolis antarctica') |> filter(!is.na(Order)) |> separate(Order, into=c('Order', 'Order_score'), sep=':') |> count(Order)
in_posi <- amph_pos_tax_cat |> filter(Sample_species == 'Posidonia australis') |> filter(!is.na(Order)) |> separate(Order, into=c('Order', 'Order_score'), sep=':') |> count(Order)
```

Unique orders in Amphibolis:
```{r}
in_amph |> nrow()
```

Unique orders in Posidonia:
```{r}
in_posi |> nrow()
```

In Amphibolis, not in Posidonia:
```{r}
setdiff(in_amph$Order, in_posi$Order) |> length()
```
```{r}
in_amph |> filter(Order %in% setdiff(in_amph$Order, in_posi$Order)) |> arrange(desc(n)) |> head(5)
```

In Posidonia, not in Amphibolis:

```{r}
setdiff(in_posi$Order, in_amph$Order) |> length()
```
```{r}
in_posi |> filter(Order %in% setdiff(in_posi$Order, in_amph$Order)) |> arrange(desc(n))
```

@@

Let's write out the table for Supplementary:
```{r}
amph_pos_tax_cat |> full_join(amph_pos_tax, by =c('sample', 'MAG')) |> select(-c(Sample_replicate.y, Sample_species.y, Tissue.y, TotalMAGs.y)) |> writexl::write_xlsx('./output/seagrass_metagenome_taxonomy.xlsx')
```
